---
title: "Lab 4: MATH 4753"
author: "Eid Zaben"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Task 1

```{r}
getwd()
```

# Task 2

```{r}
spruce.df = read.csv("SPRUCE.csv")
tail(spruce.df)
```

# Task 3

```{r}
library(s20x)
with (spruce.df, 
  trendscatter(Height~BHDiameter,f=0.5)
)

spruce.lm=with(spruce.df,lm(Height~BHDiameter)) 
height.res = residuals(spruce.lm)
print("height.res")
height.res

print("height.fit")
height.fit=fitted(spruce.lm)
height.fit

plot(height.fit,height.res)

trendscatter( height.fit,height.res)
```

The shape seen in the plot of comparing the fitted and residual values of height is a concave down quadratic graph that increases from 12 to 18 and decreases from 18 to 24. Compared to 

## TODO Finish the comparison with the first trendscatter

```{r}
plot(spruce.lm, which =1)
normcheck(spruce.lm,shapiro.wilk = TRUE)
```

The p value of the Shapiro-Wilk test is 0.29.

## TODO Null Hypothesis

The null hypothesis in this case is


# Task 4

## TODO quad.lm looks like its plotted wrong

```{r}
quad.lm=lm(Height~BHDiameter + I(BHDiameter^2),data=spruce.df)
# with(spruce.df, plot(Height~BHDiameter))
# abline(quad.lm)
plot(Height~BHDiameter,bg="Blue",pch=21,cex=1.2,
ylim=c(0,max(Height)),xlim=c(0,max(BHDiameter)), 
main="Spruce height prediction",data=spruce.df)
abline(quad.lm)
```
